\bfseries Don't interpolate your data!

### 2022-10-08 main.tex
notes to self; start the model specification

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}

@@

\paragraph{Abstract}
When there are many observations of an astronomical source---many images with different dithers, or many spectra taken at different barycentric velocities---it is often tempting to shift and stack the data, to (for example) make a high signal-to-noise average image or mean spectrum.
Bound-saturating measurements are made not by manipulating data, but instead by optimizing (or otherwise using) 
a likelihood function, where the data are treated as fixed, and model parameters are modified to fit the data.
Traditional shifting and stacking of data can be converted into a model-fitting procedure, such that no data are ever harmed.
The key component of this conversion is a spectral model that is a continuous function of wavelength (or position in the case of imaging) that can represent the psf- and pixel-convolved signal being measured by the device after any reasonable translation.
The benefits of a modeling approach are myriad:
The sacred and expensive data never have to be interpolated or otherwise modified.
Noise maps, data gaps, and bad-data masks, none of which can easily be interpolated, don't have to be manipulated either.
The simplest model-fitting procedure, like the original shift-and-add procedure, is linear in the data, so noise propagation is straightforward.
The only cost is a small increase in computational complexity.
We demonstrate all these things with toy data; we provide open-source sample code for re-use.

@@

Hello World!

@@

@@

>>>>

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

@@

\paragraph{Abstract}
When there are many observations of an astronomical source---many images with different dithers, or many spectra taken at different barycentric velocities---it is often tempting to shift and stack the data, to (for example) make a high signal-to-noise average image or mean spectrum.
Bound-saturating measurements are made not by manipulating data, but instead by manipulating a likelihood function, where the data are treated as fixed, and model parameters are modified to fit the data.
Traditional shifting and stacking of data can be converted into a model-fitting procedure, such that no data are ever harmed, and yet the output is the shift-adjusted mean spectrum.
The key component of this conversion is a spectral model that is completely flexible but also a continuous function of wavelength (or position in the case of imaging) that can represent any signal being measured by the device after any reasonable translation.
The benefits of a modeling approach are myriad:
The sacred data never have to be interpolated or otherwise modified.
Noise maps, data gaps, and bad-data masks, none of which can easily be interpolated, don't have to be manipulated.
The procedure is linear in the data, so noise propagation is straightforward.
The output can take the form of a spectrum evaluated on a pixel grid, as is expected by the spectroscopic community.
The only cost is a small increase in computational complexity.
We demonstrate all these things with both toy and real data; we provide open-source sample code for re-use.

@@

Things to emphasize here and below:
\begin{itemize}
\item The input--output relationship of this modeling approach is identical to the input--output relationship of the shift-and-coadd approach. It is just better.
\item People have a lot of hangups about the word ``model''. We have to use another word, or be very very careful how we use this word.
\item We have to explain why and how we are better than other approaches in terms that people care about (like radial-velocity precision and abundances etc).
\item We should choose a rigid terminology and stick to it here. HOGG is currently liking ``combined spectrum'' for the model expectation.
\end{itemize}

@@

The problem set-up is that there are $N$ spectral measurements $y_i$ (with $1\leq i\leq N$), each of which is a $M_i$-pixel list of intensities (fluxes maybe).
Different observations $i$ might have different numbers $M_i$ of pixels.
The $M_i$ pixels are associated with an $M_i$-length list of positions (wavelengths or log-wavelengths) $x_i$.
Each measurement or observation $i$ is shifted by a shift operator $\Delta x_i$, which can be thought of as a length-$M_i$ list of pixel offsets or shifts, such that the ``rest frame'' or ``unshifted'' positions corresponding to the length-$M_i$ list of data $y_i$ is the list of differences $x_i - \Delta x_i$.
The goal is to get a combined spectrum that represents the mean of these spectra, in the common rest frame.

The spectral expectation model $f(x;\theta)$, which will be a representation for our combined spectrum, is a function of $x$ controlled by a list $\theta$ of $P$ linear parameters $a_j$ (with $1\leq j\leq P$).
This function will be very flexible; possibly even an interpolation of control points or a Fourier series; the number $P$ will be roughly the number of pixels we want in our output combined spectrum.
The statistical model underlying can be expressed qualitatively as
\begin{align}
    y_i &= f(x_i + \Delta x_i;\theta) + \mbox{noise} \\
    f(x;\theta) &= \sum_{j=1}^P a_j\,g_j(x) ~,
\end{align}
where we are being a bit loose with terminology (since $x_i$ is a set of positions, not just one position), the noise is additive (but not yet specified), and each of the $P$ functions $g_j(x)$ is a basis function of the representation.

HOGG: NOISE MODEL

HOGG: BASIS FUNCTIONS.

@@

\paragraph{Acknowledgements}
It is a pleasure to thank
Adrian Price-Whelan (Flatiron),
and the Astronomical Data Group at the Flatiron Institute
for valuable discussions.
